Abstract

### 2025-02-15 output/Letter_to_editor_reviewers.Rmd
Finish adding (initial) responses to reviewer 3

We apologize for the lack of clarity.
We intended to refer to footnote 1 in page 16, we have correctly referenced the page and footnote in the new version.

@@

@@

@@

>>>>

We apologize for the lack of clarity.
We intended to refer to footnote 1 in page 16, we have now completed the reference.

@@

That is exactly the case.
All graphical devices require that coordinates are represented in a rectangular Cartesian system such that the first coordinate maps to the horizontal dimension and the second coordinate to the vertical one.
Therefore, even when computing the coordinates in a non-orthonormal basis, we need to transform them to the canonical one to plot them.
This is explained in our manuscript, in p. 19 (lines 9-14) in the original version, and in p. <!-- # TODO: Complete --> (lines <!-- # TODO: Complete -->) in the new submission.
We are unsure about how to make this clearer, we are afraid, so we would appreciate any suggestions.

@@

<!-- # NOTE: Help needed -->

@@

To our understanding, the representation is always sample dependent, because item parameters are in turn.
Our argument is that the non-orthogonal coordinate system is a more faithful representation of the latent covariance structure, as the cosines between the axes are the geometric equivalent of the latent trait correlations.
This is, to our understanding, the value of using this more complex representation, and so we have tried to convey in (please see p. 31, line 19 to p. 32, line 4, in the original version.

We acknowledge though that this may have been explained without enough clarity, and therefore strived to improve this discussion.
The aforementioned paragraph now reads:

    > [...] This implies that the axes defined by the test space basis,
    and not the ones defined by the latent space basis,
    are a geometrical representation of the covariance structure:
    Indeed, the cosines of the angles among the test space axes are
    the geometric equivalent of the correlations. [...]

### 2025-02-15 output/Main_text.Rmd
Finish adding (initial) responses to reviewer 3

On the other hand, establishing the relationship between the original and
the orthonormalized basis has provided us with an important insight:
The distinction between the latent space and the test space.
Our derivations show that, to properly represent the items,
we need to define a vector space that has
the latent covariance (or correlation) matrix as its inner product matrix.
On the other hand, the latent space,
where the person parameters are represented,
has the inverse of the covariance (correlation) matrix as
its inner product matrix.
This implies that the axes defined by the test space basis,
and not the latent space basis,
are a geometrical representation of the covariance structure.
This result is both surprising and counterintuitive
---Several instances in the MIRT literature suggest that
"the angle between the axes [of the latent ability space] represents
the degree of correlation" [@ackerman_multidimensional_2005-1, p. 16].
We hope our results help shed a clearer light on
this often misunderstood topic.

>>>>

On the other hand, establishing the relationship between the original and
the orthonormalized basis has provided us with an important insight:
The distinction between the latent space and the test space.
Our derivations show that, to properly represent the items,
we need to define a vector space that has
the latent covariance (or correlation) matrix as its inner product matrix.
On the other hand, the latent space,
where the person parameters are represented,
has the inverse of the covariance (correlation) matrix as
its inner product matrix.
This implies that the axes defined by the test space basis,
and not the ones defined by the latent space basis,
are a geometrical representation of the covariance structure:
Indeed, the cosines of the angles among the test space axes are
the geometric equivalent of the correlations.
This result is both surprising and counterintuitive
---Several instances in the MIRT literature suggest that
"the angle between the axes [of the latent ability space] represents
the degree of correlation" [@ackerman_multidimensional_2005-1, p. 16].
We hope our results help shed a clearer light on
this often misunderstood topic.
